Equalizer Programming Guide

### 2009-01-10 Developer/ProgrammingGuide/paper.tex
Programming Guide: Add compositor documentation
[ ] May break build
[ ] Breaks existing applications
[ ] Bugfix
[ ] New Feature
[ ] Cleanup
[ ] Optimization
[x] Documentation

\date{
  \vspace{15cm}
  {\Large\sffamily The official reference for creating parallel,
    scalable OpenGL\texttrademark\ applications with the Equalizer
    parallel rendering framework}\\\vspace{1cm}
  Version 1.5.1 for Equalizer 0.6+svn\\[\medskipamount]
%  \today
}

@@

\textcopyright 2007-2008 \htmladdnormallink{Eyescale Software GmbH}
{http://www.eyescale.ch}. All rights reserved. No permission is
granted to copy, distribute, or create derivative works from the
contents of this electronic documentation in any manner, in whole or in
part, without the prior written permission of Eyescale Software GmbH.

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
\listoffigures
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.5.1   & Dec 08, 2008 & Minor updates based on the print version\\
    \hline
  \end{tabularx}}
\clearpage

@@

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/slices.pdf}
  {\caption{\small\label{fSlices}Volume Rendering using 3D Textures}}
\end{wrapfigure}
When using 2D (sort-first) or stereo decompositions, no special
programming is needed to achieve good scalability, as \textsf{eVolve} is
mostly fill-limited and therefore scales nicely in these modes.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/volResult.png}
  {\caption{\small\label{fVolResult}Result of \fig{fBlendPerspective}}}
\end{wrapfigure}
For recomposition, the 2D frame buffer contents are blended to form a
seamless picture. For correct blending, the frames are ordered in the
same back-to-front order as the slices used for rendering, and use the
same blending parameters. Simplified, the frame buffer images are
`thick' slices which are `rendered' by writing their content to the
destination frame buffer using the correct order.

@@

Finding the correct assembly order for perspective frusta is more
complex. The perspective distortion invalidates a simple orientation
criteria like the one used for orthographic frusta. For the view and
frustum setup shown in \fig{fBlendPerspective}\footnote{Volume Data Set
  courtesy of: AVS, USA} the correct compositing order is 4-3-1-2 or
1-4-3-2.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/statistics.png}
  {\caption{\small\label{fStatistics}Statistics for a two-node 2D compound}}
\end{wrapfigure}
Statistics are enabled per entity using an attribute hint. The hint
determines how precise the gathered statistics are. When set to
\textsf{fastest}, the per-frame clock is sampled directly when the event
occurs. When set to \textsf{nicest}, all OpenGL commands will be
finished before sampling the event. This incurs a performance penalty,
but gives more correct results. The default setting is fastest in
release builds, and nicest in debug builds.

The events are processed by the channel's and window's
\textsf{processEvent} method. The default implementation sends these
events to the config, using \textsf{Config::sendEvent}, as explained in
\sref{sEventHandling}. When the default implementation of
\textsf{Config::handleEvent} receives the statistics event, it sorts the
event per frame and per originator. When a frame has been finished, the
events are pushed to the local (app-)node for visualization.

>>>>

\date{
  \vspace{15cm}
  {\Large\sffamily The official reference for creating parallel,
    scalable OpenGL\texttrademark\ applications with the Equalizer
    parallel rendering framework}\\\vspace{1cm}
  Version 1.5.2 for Equalizer 0.6+svn\\[\medskipamount]
%  \today
}

@@

\textcopyright 2007-2009 \htmladdnormallink{Eyescale Software GmbH}
{http://www.eyescale.ch}. All rights reserved. No permission is
granted to copy, distribute, or create derivative works from the
contents of this electronic documentation in any manner, in whole or in
part, without the prior written permission of Eyescale Software GmbH.

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
\listoffigures
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.5.1   & Dec 08, 2008 & Minor updates based on the print version\\
    1.5.2   & Jan 10, 2009 & Added \sref{sCompositor}\\
    \hline
  \end{tabularx}}
\clearpage

@@

\subsubsection{\label{sCompositor}The Compositor}

%\begin{wrapfigure}{r}{.618\textwidth}
\begin{figure}[h!t]\center
  \includegraphics[width=.9\textwidth]{images/compositor.pdf}
  {\caption{\small\label{fCompositor}Functional diagram of the compositor}}
\end{figure}
The \textsf{Compositor} class gathers a set of static functions which
implement the various compositing algorithms and low-level
optimizations. \fig{fCompositor} provides a top-down functional overview
of the various compositor functions.

On a high level, the compositor combines multiple input frames using 2D
tiling, depth-compositing for polygonal data or sorted, alpha-blended
compositing for semi-transparent volumetric data. These operations
composite either directly all images on the GPU, or use a CPU-based
compositor and then transfer the pre-integrated result to the GPU. The
high-level entry points automatically select the best algorithm. The
CPU-based compositor uses, based on availability, OpenMP and the
Paracomp compositing library to accelerate its operation.

On the next lower level, the compositor provides functionality to
composite a single frame, either using 2D tiling (possibly with blending
for alpha-blended compositing) or depth-based compositing.

The per-frame compositing in turn relies on the per-image compositing
functionality, which automatically decides on the algorithm to be used
(2D or depth-based). The concrete per-image assembly operation uses
OpenGL operations to composite the pixel data into the framebuffer,
potentially using GLSL for better performance.

@@

When using 2D (sort-first) or stereo decompositions, no special
programming is needed to achieve good scalability, as \textsf{eVolve} is
mostly fill-limited and therefore scales nicely in these modes.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/slices.pdf}
  {\caption{\small\label{fSlices}Volume Rendering using 3D Textures}}
\end{wrapfigure}
For recomposition, the 2D frame buffer contents are blended to form a
seamless picture. For correct blending, the frames are ordered in the
same back-to-front order as the slices used for rendering, and use the
same blending parameters. Simplified, the frame buffer images are
`thick' slices which are `rendered' by writing their content to the
destination frame buffer using the correct order.

@@

\begin{wrapfigure}{l}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/volResult.png}
  {\caption{\small\label{fVolResult}Result of \fig{fBlendPerspective}}}
\end{wrapfigure}
Finding the correct assembly order for perspective frusta is more
complex. The perspective distortion invalidates a simple orientation
criteria like the one used for orthographic frusta. For the view and
frustum setup shown in \fig{fBlendPerspective}\footnote{Volume Data Set
  courtesy of: AVS, USA} the correct compositing order is 4-3-1-2 or
1-4-3-2.

@@

Statistics are enabled per entity using an attribute hint. The hint
determines how precise the gathered statistics are. When set to
\textsf{fastest}, the per-frame clock is sampled directly when the event
occurs. When set to \textsf{nicest}, all OpenGL commands will be
finished before sampling the event. This incurs a performance penalty,
but gives more correct results. The default setting is fastest in
release builds, and nicest in debug builds.

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/statistics.png}
  {\caption{\small\label{fStatistics}Statistics for a two-node 2D compound}}
\end{wrapfigure}
The events are processed by the channel's and window's
\textsf{processEvent} method. The default implementation sends these
events to the config, using \textsf{Config::sendEvent}, as explained in
\sref{sEventHandling}. When the default implementation of
\textsf{Config::handleEvent} receives the statistics event, it sorts the
event per frame and per originator. When a frame has been finished, the
events are pushed to the local (app-)node for visualization.
